---
title: "Corona Virus Spread Analysis"
author: "Anish Singh Walia"
date: "2/11/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## AIM

The primary objective and aim of this project is to analyze the spread of Corona Virus worldwide. This dataset has data available from 22 Jan,2020 till 9th Feb 2020. The csv file is taken from WHO(World Health Organization).

### What is Corona Virus(nCOV)

Novel Coronavirus (nCoV) is a virus (more specifically, a coronavirus) identified as the cause of an outbreak of respiratory illness first detected in Wuhan, China. Early on, many of the patients in the outbreak in Wuhan, China reportedly had some link to a large seafood and animal market, suggesting animal-to-person spread. However, a growing number of patients reportedly have not had exposure to animal markets, indicating person-to-person spread is occurring. At this time, it’s unclear how easily or sustainably this virus is spreading between people




#### Let's begin with importing the data

```{r message=FALSE,warning=FALSE}

require(dplyr)
require(highcharter)
require(ggplot2)
require(readr)

nCOV<-read_csv("data/2019_nCoV_data.csv",col_names = TRUE)

#ommitting any NA values if there
nCOV<-na.omit(nCOV)
#attaching the data frame
attach(nCOV)


```


#### Total deaths between that timeframe

```{r message=FALSE,warning=FALSE}

summary(nCOV)


sum(Deaths) #total deaths of 6126 confirmed between 22 Jan -  9th Feb

#Let's find sum of confirmed cases: 286573 people affected with nCOV worldwide between 22 Jan - 9 Feb
sum(Confirmed)

# let's see how many have recovered : out of 286573, only 13455 have recovered from it between 22 Jan - 9th Feb
sum(Recovered)

```

Let's make a data frame out of the above totals to understand better.

```{r message=FALSE,warning=FALSE}

ncov_cases<-rbind(sum(Deaths),sum(Confirmed),sum(Recovered)) #combining by rows
ncov_cases <- as.data.frame(ncov_cases) #converting it to a data frame

#changing the names for better understanding
row.names(ncov_cases)<-c("Death","Confirmed","Recovered")
names(ncov_cases)[1]<-"Count"

ncov_cases


#What is the recovery rate
recovery_rate<-(ncov_cases$Count[3]/ncov_cases$Count[2])*100 
recovery_rate
#hence, the recovery rate is pretty low i.e 4.69%

death_rate<-(100-recovery_rate)
death_rate




```

So we can notice that the recovery rate of perople affected with nCOV virus is approxinately __4.7 %__, i.e only 4.7 % of people disgnosed with nCOV __recovered__ and the rest __died__. Hence the death rate is __95.3 %__ for nCOV.




#### Let's do some country wise analysis


Let's first find out number of confirmed cases country wise. Obviously __Mainland China__ is going to have the maximum number as it is the Epicenter of the breakout.


```{r warning=FALSE,message=FALSE}

#let's check the number of nCOV cases per country
count_country<-nCOV %>% 
  group_by(Country) %>%
  select(Confirmed) %>% 
  summarise(nCases=sum(Confirmed))


  
#Mainland China has aroung 284905 confirmed cases out of total 286573 nCOV csases and this makes sense.
   
#let's plot these, and lets ingnore China 

count_country1<-count_country %>% 
  filter(count_country$Country != "Mainland China") %>% 
  arrange(desc(count_country1$nCases))
        

hchart(count_country1 , type = "column", hcaes(x=count_country1$Country, y= count_country1$nCases ),color="purple") %>% 
      hc_exporting(enabled = TRUE) %>%
      hc_title(text="Confirmed Cases grouped by Country",align="center") %>%
      hc_add_theme(hc_theme_elementary()) 



```

In in the above data frame we can notice that __Mainland China__ has around 284905 confirmed cases out of total 286573 nCOV csases and this makes sense. That makes Mainland China having 99.4 % of the total confirmed cases.


#### Let's find the number of Deaths occured in these countries.


```{r warning=FALSE,message=FALSE}

count_deaths<-nCOV %>% 
  group_by(Country) %>% 
  select(Deaths) %>% 
  summarise(nDeaths=sum(Deaths)) %>% 
  arrange(desc(nDeaths))
  

count_deaths
```

*__As per this dataset between 22 Jan till 9th Feb all the deaths due to nCOV have only been in Mainland China and Hong Kong.__*


