---
title: "Corona Virus Spread Analysis"
author: "Anish Singh Walia"
date: "2/11/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## AIM

The primary objective and aim of this project is to analyze the spread of Corona Virus worldwide. This dataset has data available from 22 Jan,2020 till 9th Feb 2020. The csv file is taken from WHO(World Health Organization).

### What is Corona Virus(nCOV)

Novel Coronavirus (nCoV) is a virus (more specifically, a coronavirus) identified as the cause of an outbreak of respiratory illness first detected in Wuhan, China. Early on, many of the patients in the outbreak in Wuhan, China reportedly had some link to a large seafood and animal market, suggesting animal-to-person spread. However, a growing number of patients reportedly have not had exposure to animal markets, indicating person-to-person spread is occurring. At this time, it’s unclear how easily or sustainably this virus is spreading between people




#### Let's begin with importing the data

```{r message=FALSE,warning=FALSE}

require(dplyr)
require(highcharter)
require(ggplot2)
require(readr)

nCOV<-read_csv("data/2019_nCoV_data.csv",col_names = TRUE)

#ommitting any NA values if there
nCOV<-na.omit(nCOV)
#attaching the data frame
attach(nCOV)


```


#### Total deaths between that timeframe

```{r message=FALSE,warning=FALSE}

summary(nCOV)


sum(Deaths) #total deaths of 6126 confirmed between 22 Jan -  9th Feb

#Let's find sum of confirmed cases: 286573 people affected with nCOV worldwide between 22 Jan - 9 Feb
sum(Confirmed)

# let's see how many have recovered : out of 286573, only 13455 have recovered from it between 22 Jan - 9th Feb
sum(Recovered)

```

